---
title: "Functionality of fuzzyclara package"
author: "Maximilian Weigert, Asmik Nalmpatian, Jana Gauss, Alexander Bauer"
date: "`r format(Sys.time(), '%d.%B %Y')`"
output:
  html_document:
    toc: yes
  vignette: >
    %\VignetteIndexEntry{Functionality of TouristClust package}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
  rmarkdown::html_vignette: default
editor_options:
  chunk_output_type: console
---

This document gives an overview of the functionality provided by the R package
`fuzzyclara`.
All examples are based on the dataset `USArrests` which lists four
criminal statistics for each federal state in the U.S. in 1973:

- column `Murder` contains the number of murder arrests per 100,000 inhabitants
- column `Assault` contains the number of assault arrests per 100,000 inhabitants
- column `UrbanPop` contains the percentage of the population living in urban areas
- column `Rape` contains the number of rape arrests per 100,000 inhabitants

Exemplarily, let's take a quick look at the first rows of `USArrests`:

```{r}
head(USArrests)
```

Based on these four dimensions we will try to cluster the federal states
into homogeneous groups throughout all of the examples in this vignette.

For starting the analyses, let's first load the `fuzzyclara` package,
and additionally the `dplyr` package for general data handling.

```{r packages, message = FALSE}
library(fuzzyclara)
library(dplyr)
```


## Clustering

### Hard clustering

Applying function `fuzzyclara` with argument `type = "hard"` allows us to perform a
hard clustering of the federal states utilizing the CLARA clustering algorithm.
In this example, let's compute a cluster solution with 3 clusters based on the
euclidean distance metric.

```{r, hard clustering example}
cc_hard <- fuzzyclara(data        = USArrests,
                      clusters    = 3,
                      metric      = "euclidean",
                      type        = "hard",
                      seed        = 3526,
                      verbose     = 0)
```

The output object contains the `Medoids` -- i.e., the most central / most *typical* observation
for each of the three clusters --, the `Clustering` vector specifying for each observation
which cluster it belongs to, and the `Minimum average distance`, which is defined as the average distance between each observation to its cluster, with `Minimum` referring to
this average distance value of the outputted optimal cluster solution.

```{r}
cc_hard
```

### Fuzzy clustering

Alternatively, it is possible to perform a fuzzy clustering by setting
`type = "fuzzy"`. Additionally, the *fuzziness parameter* `m` could be set to
any value between 1 and 2. 2, i.e. the default value, encodes full fuzzyness, and
1 encodes no fuzziness at all, i.e. a hard clustering.

```{r, fuzzy clustering example}
cc_fuzzy <- fuzzyclara(data        = USArrests,
                       clusters    = 3,
                       metric      = "euclidean",
                       type        = "fuzzy",
                       m           = 2,
                       seed        = 3526,
                       verbose     = 0)
```

The output contains the same elements as the hard clustering output above.
Additionally it includes a matrix of the `Membership scores`. For each federal state
this matrix contains the membership probabilities to the three different clusters.

```{r}
cc_fuzzy
```


### Clustering with self-defined distance function and other distance functions

It is possible to pass your own distance function to function `fuzzyclara`
through the argument `metric`, as shown in the following example and as explained
in `?fuzzyclara`.

```{r, self-defined distance function, warning = FALSE}
# define a quadratic distance function
dist_function <- function(x, y) {
  sqrt(sum((x - y)^2))
}

cc_dist <- fuzzyclara(data        = USArrests,
                      clusters    = 3,
                      metric      = dist_function,
                      type        = "fuzzy",
                      m           = 2,
                      seed        = 3526,
                      verbose     = 0)
cc_dist
```

You can also use other distance functions from the `proxy` package like Gower, Manhattan etc. In order to specify arguments of the distance metric (e. g. *p* for Minkowski distance), you can use a self-defined distance function.

```{r manhattan, warning = FALSE}
# Example 1: Manhattan distance
cc_manh <- fuzzyclara(data        = USArrests,
                      clusters    = 3,
                      metric      = "manhattan",
                      samples     = 1,
                      sample_size = NULL,
                      type        = "hard",
                      seed        = 3526,
                      verbose     = 0)
cc_manh

# Example 2: Minkowski distance with parameter 'p = 1'
dist_mink <- function(x, y) {
  proxy::dist(list(x, y), method = "minkowski", p = 1)
}
cc_mink <- fuzzyclara(data        = USArrests,
                      clusters    = 3,
                      metric      = dist_mink,
                      samples     = 1,
                      sample_size = NULL,
                      type        = "hard",
                      seed        = 3526,
                      verbose     = 0)
cc_mink
```


### Select optimal number of clusters

To evaluate the optimal number of clusters for your application, you can 
define a to be evaluated `clusters_range` in function `evaluate_cluster_numbers`.
The function produces a scree plot to compare the cluster solutions.

```{r optimal number clusters, warning = FALSE}
cc_number <- evaluate_cluster_numbers(
  data            = USArrests,
  clusters_range  = 2:6,
  metric          = "euclidean",
  type            = "hard",
  seed            = 3526,
  verbose         = 0
)

cc_number
```


## Plot of clustering results

This section showcases all types of visualizations implemented in `fuzzyclara`.
To have some more exemplary data to plot, let's first add the `State` (i.e., the 
name of the federal state) and the `Area` (i.e., West / Midwest / South / Northeast)
information as proper columns to the data.

```{r, enrichment example}
# define state areas
states_west      <- c("Washington", "Oregon", "California", "Nevada", "Arizona",
                      "Idaho", "Montana", "Wyoming", "Colorado", "New Mexico",
                      "Utah", "Hawaii", "Alaska")
states_south     <- c("Texas", "Oklahoma", "Arkansas", "Louisiana", "Mississippi",
                      "Alabama", "Tennessee", "Kentucky", "Georgia", "Florida",
                      "South Carolina", "North Carolina", "Virginia", "West Virginia")
states_midwest   <- c("Kansas", "Nebraska", "South Dakota", "North Dakota",
                      "Minnesota", "Missouri", "Iowa", "Illinois", "Indiana",
                      "Michigan", "Wisconsin", "Ohio")
states_northeast <- c("Maine", "New Hampshire", "New York", "Massachusetts",
                      "Rhode Island", "Vermont", "Pennsylvania", "New Jersey",
                      "Connecticut", "Delaware", "Maryland")

# enrich data with 'State' and 'Area' information as proper columns
USArrests_enriched <- USArrests %>% 
  mutate(State = rownames(USArrests),
         State = factor(State),
         Area  = case_when(State %in% states_west      ~ "West",
                           State %in% states_south     ~ "South",
                           State %in% states_midwest   ~ "Midwest",
                           State %in% states_northeast ~ "Northeast"),
         Area  = factor(Area))
```

### Boxplot

Simply applying the general `plot` function to an object of class `fuzzyclara`
we can use boxplots to visualize the differences between the clusters regarding
a metric variable.
In this example, let's compare the distribution of the `Assault` variable between
the clusters obtained using the hard clustering approach.

```{r}
plot(x        = cc_hard,
     data     = USArrests_enriched,
     variable = "Assault") 
```

### Barplot

If the variable that should be plotted is not metric, but categorical,
the function defaults to displaying a stacked bar plot.

```{r}
plot(x        = cc_hard,
     data     = USArrests_enriched,
     variable = "Area")
```

### Wordcloud

Specifying `type = "wordclouds"` as an argument to the `plot` function,
it is possible to draw one wordcloud per cluster based on a specific variable.
Due to the lack of a proper character variable, we here examplarily apply the
wordcloud to the `State` variable. Please note, though, that this is obviously
not the standard application for a wordcloud due to every state name appearing
only once in the dataset.

```{r fig.width=6, fig.height=4}
plot(x        = cc_hard,
     data     = USArrests_enriched,
     variable = "State",
     type     = "wordclouds")
```

### Scatterplot

Instead of only displaying the distribution of one single variable,
it is also possible to specify `type = "scatterplot"` and both the arguments
`x_var` and `y_var`, to create a scatterplot of the data, joint with
estimated cluster-specific linear trends.

```{r, message=FALSE}
plot(x     = cc_hard,
     data  = USArrests_enriched,
     type  = "scatterplot",
     x_var = "Murder",
     y_var = "Assault")
```

For a fuzzy clustering solution, you can additionally enrich the scatterplot
by each observation's membership probability, plotted as each point's transparency
level. Simply specify `focus = TRUE`:

```{r}
# plot of all clusters
plot(x     = cc_fuzzy,
     data  = USArrests_enriched,
     type  = "scatterplot",
     x_var = "Murder",
     y_var = "Assault", 
     focus = TRUE)

# plot only of selected clusters, by specifying argument 'focus_clusters'
plot(x              = cc_fuzzy,
     data           = USArrests_enriched,
     type           = "scatterplot",
     x_var          = "Murder",
     y_var          = "Assault", 
     focus          = TRUE,
     focus_clusters = 1)
```


### PCA

The cluster solution can be visualized based on a Principal Component Analysis (PCA)
by specifying `type = "pca"` in the base `plot` function.
Additionally, argument `group_by` can be used to specify a variable which should
be depicted as different point shapes.

```{r fig.width=6, fig.height=4}
plot(x        = cc_hard,
     data     = USArrests_enriched, 
     type     = "pca",
     group_by = "Area")
```

For fuzzy clustering, you can focus on one or more clusters and also plot the membership probability.

```{r}
# plot of all clusters
plot(x     = cc_fuzzy,
     data  = USArrests_enriched,
     type  = "pca",
     focus = TRUE)

# plot only of selected clusters, by specifying argument 'focus_clusters'
plot(x              = cc_fuzzy,
     data           = USArrests_enriched,
     type           = "pca",
     focus          = TRUE,
     focus_clusters = 1)
```


### Silhouette

A standard silhouette plot can be plotted by setting `type = "silhouette"`.
The output object of the base `plot` function will be a list, including the visualization
as element `plot`, the per-cluster average silhouette value as element `silhouette_table`
and the overall average silhouette value as element `average_silhouette_width`.

```{r fig.width=9, fig.height=4}
silhouettes <- plot(x    = cc_hard,
                    data = USArrests,
                    type = "silhouette")

silhouettes$plot

silhouettes$silhouette_table

silhouettes$average_silhouette_width
```

### Fuzzy clustering: Threshold for membership scores

As an additional option for the PCA plot for fuzzy clustering solutions,
it is possible to set the `memebership_threshold` argument.
All observations with a membership probability above this threshold will be
recognized as being part of the respective *core cluster* and will be highlighted accordingly
in the PCA plot.

If we set `membership_threshold = 0` all observations belonging to one cluster
are part of its core cluster.

```{r fig.width=6, fig.height=4}
plot(x                    = cc_fuzzy,
     data                 = USArrests_enriched,
     type                 = "pca",
     variable             = "Assault",
     membership_threshold = 0)
```

If we instead set `membership_threshold = 0.5`, only the few observations
with a membership probability above 0.5 will be highlighted as being part of the
core cluster.

```{r fig.width=6, fig.height=4}
plot(x                    = cc_fuzzy,
     data                 = USArrests_enriched,
     type                 = "pca",
     variable             = "Assault",
     membership_threshold = 0.5)
```

The argument `membership_threshold` has a similar effect on scatterplots.

```{r fig.width=6, fig.height=4, message = FALSE}
# scatterplot with membership_threshold 0
plot(x                    = cc_fuzzy,
     data                 = USArrests_enriched,
     type                 = "scatterplot",
     x_var                = "Murder",
     y_var                = "Assault",
     membership_threshold = 0)

# scatterplot with membership_threshold 0.5
plot(x                    = cc_fuzzy,
     data                 = USArrests_enriched,
     type                 = "scatterplot",
     x_var                = "Murder",
     y_var                = "Assault",
     membership_threshold = 0.5)
```
