---
title: "Functionality of TouristClust package (interactive)"
author: "Maximilian Weigert, Asmik Nalmpatian, Alexander Bauer"
date: "`r format(Sys.time(), '%d.%B %Y')`"
output:
  html_document:
    toc: yes
  vignette: >
    %\VignetteIndexEntry{Functionality of TouristClust package (interactive)}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
  rmarkdown::html_vignette: default
editor_options:
  chunk_output_type: console
runtime: shiny
---

This document gives an overview of the functionality provided by the R package
`TouristClust`.

```{r packages, message = FALSE}
library(TouristClust)
library(dplyr)
library(purrr)
library(shiny)
library(rlang)
```


## Clustering

### Hard clustering
```{r, hard clustering example}
# TODO replace USArrests by a better dataset (maybe TourIST)
cc_fixed <- claraclust(data        = USArrests,
                       clusters    = 3,
                       metric      = "euclidean",
                       samples     = 1,
                       sample_size = NULL,
                       type        = "fixed",
                       seed        = 3526,
                       verbose     = 0)
cc_fixed
```

### Fuzzy clustering
```{r, fuzzy clustering example}
cc_fuzzy <- claraclust(data        = USArrests,
                       clusters    = 3,
                       metric      = "euclidean",
                       samples     = 1,
                       sample_size = NULL,
                       type        = "fuzzy",
                       m           = 2,
                       seed        = 3526,
                       verbose     = 0)
cc_fuzzy
```

### Clustering with self-defined distance function and other distance functions
```{r, self-defined distance function}
dist_function <- function(x, y) {
  sqrt(sum((x - y)^2))
}

cc_dist <- claraclust(data        = USArrests,
                      clusters    = 3,
                      metric      = dist_function,
                      samples     = 1,
                      sample_size = NULL,
                      type        = "fuzzy",
                      m           = 2,
                      seed        = 3526,
                      verbose     = 0)
cc_dist
```

You can also use other distance functions from the proxy package like Gower, Manhattan etc. In order to specify arguments of the distance metric (e. g. *p* for Minkowski distance), you can use a self-defined distance function.
```{r manhattan, warning = FALSE}
cc_manh <- claraclust(data        = USArrests,
                       clusters    = 3,
                       metric      = "manhattan",
                       samples     = 1,
                       sample_size = NULL,
                       type        = "fixed",
                       seed        = 3526,
                       verbose     = 0)
cc_manh

dist_mink <- function(x, y) {
  proxy::dist(list(x, y), method = "minkowski", p = 1)
}
cc_mink <- claraclust(data        = USArrests,
                       clusters    = 3,
                       metric      = dist_mink,
                       samples     = 1,
                       sample_size = NULL,
                       type        = "fixed",
                       seed        = 3526,
                       verbose     = 0)
cc_mink
```

### Select optimal number of clusters
```{r optimal number clusters, warning = FALSE, fig.width=6, fig.height=4}
cc_number <- plot_cluster_numbers(
                       data            = USArrests,
                       clusters_range  = 2:6,
                       metric          = "euclidean",
                       samples         = 1,
                       sample_size     = NULL,
                       type            = "fixed",
                       seed            = 3526,
                       verbose         = 0)
cc_number
```

```{r optimal number clusters interactive, warning = F, echo = F}


shinyApp(
  ui = fluidPage(
    numericInput("num_min", 
                 label = "Cluster number minimum",min = 2, max = 7, value = 2, step = 1),
    numericInput("num_max", 
                 label = "Cluster number maximum",min = 3, max = 8, value = 6, step = 1),
    plotOutput("plot_cluster_num")
  ),
  
  
  server = function(input, output) {
    
    output$plot_cluster_num <- renderPlot({
      plot_cluster_numbers(
                       data            = USArrests,
                       clusters_range  = input$num_min:input$num_max,
                       metric          = "euclidean",
                       samples         = 1,
                       sample_size     = NULL,
                       type            = "fixed",
                       seed            = 3526,
                       verbose         = 0)
    })
        
  },
  options = list(height = 800, width = 600)
)

```


## Plot of clustering results
```{r, enrichment example}
# Enrich the USArrest dataset by area and state

USArrests_enriched <- USArrests %>% 
  mutate(State = as.factor(rownames(USArrests)),
         Area  = as.factor(case_when(State %in% c("Washington", "Oregon",
                 "California", "Nevada", "Arizona", "Idaho", "Montana",
                 "Wyoming", "Colorado", "New Mexico", "Utah", "Hawaii",
                 "Alaska") ~ "West",
                 State %in% c("Texas", "Oklahoma", "Arkansas", "Louisiana",
                 "Mississippi", "Alabama", "Tennessee", "Kentucky", "Georgia",
                 "Florida", "South Carolina", "North Carolina", "Virginia",
                 "West Virginia") ~ "South",
                 State %in% c("Kansas", "Nebraska", "South Dakota",
                 "North Dakota", "Minnesota", "Missouri", "Iowa", "Illinois",
                 "Indiana", "Michigan", "Wisconsin", "Ohio") ~ "Midwest",
                 State %in% c("Maine", "New Hampshire", "New York",
                 "Massachusetts", "Rhode Island", "Vermont", "Pennsylvania",
                 "New Jersey", "Connecticut", "Delaware", "Maryland") ~
                   "Northeast")))

```

### Boxplot
```{r fig.width=6, fig.height=4}
plot(x = cc_fuzzy, data = USArrests_enriched, variable = "Assault") 
```


```{r,echo=F, message = F}
shinyApp(
  ui = fluidPage(
    sliderInput("threshold", 
                 label = "Confidence threshold",min = 0, max = 1, value = 0.5, step = 0.05),
    plotOutput("cityplt")
  ),
  
  
  server = function(input, output) {
    
    output$cityplt <- renderPlot({
      plot(
      x = cc_fuzzy,
      data = USArrests_enriched,
      variable = "Assault",
      confidence_threshold = input$threshold
      )
    })
        
  },
  options = list(height = 800, width = 600)
)

```


### Barplot
```{r fig.width=6, fig.height=4}
plot(x = cc_fuzzy, data = USArrests_enriched, variable = "Area")

```


```{r,echo=F, message = F}
shinyApp(
  ui = fluidPage(
    sliderInput("threshold", 
                 label = "Confidence threshold",min = 0, max = 1, value = 0.5, step = 0.05),
    plotOutput("cityplt")
  ),
  
  
  server = function(input, output) {
    
    output$cityplt <- renderPlot({
      plot(
      x = cc_fuzzy,
      data = USArrests_enriched,
      variable = "Area",
      confidence_threshold = input$threshold
      )
    })
        
  },
  options = list(height = 800, width = 600)
)

```

### Wordcloud
```{r fig.width=6, fig.height=4}
plot(x = cc_fuzzy, data = USArrests_enriched, variable = "State",
     type = "wordclouds")
```

### Scatterplot
```{r fig.width=6, fig.height=4}
plot(x = cc_fuzzy, data = USArrests_enriched, type = "scatterplot",
     variable = "Murder", group_by = "Assault")

```

```{r,echo=F, message = F}
shinyApp(
  ui = fluidPage(
    sliderInput("threshold", 
                 label = "Confidence threshold",min = 0, max = 1, value = 0.5, step = 0.05),
    plotOutput("cityplt")
  ),
  
  
  server = function(input, output) {
    
    output$cityplt <- renderPlot({
      plot(
      x = cc_fuzzy,
      data = USArrests_enriched,
      type = "scatterplot",
      variable = "Murder", 
      group_by = "Assault",
      confidence_threshold = input$threshold,
      plot_all_fuzzy = TRUE
      )
    })
        
  },
  options = list(height = 800, width = 600)
)

```


### PCA
```{r fig.width=6, fig.height=4}
plot(x = cc_fuzzy, data = USArrests_enriched, type = "pca",
     group_by = "Area")
```

```{r,echo=F, message = F}
shinyApp(
  ui = fluidPage(
    sliderInput("threshold", 
                 label = "Confidence threshold",min = 0, max = 1, value = 0.5, step = 0.05),
    plotOutput("cityplt")
  ),
  
  
  server = function(input, output) {
    
    output$cityplt <- renderPlot({
      plot(
      x = cc_fuzzy,
      data = USArrests_enriched,
      type = "pca",
      group_by = "Area",
      confidence_threshold = input$threshold
      )
    })
        
  },
  options = list(height = 800, width = 600)
)

```

### Silhouette
```{r fig.width=6, fig.height=4}
plot(x = cc_fuzzy, data = USArrests, type = "silhouette")
```
